Internal models of type theory

### 2013-10-24 article-models-eg.tex
putting the new macros \ctxext, \ctxwk and \subst into use

When $A$ is a type in context
$\Gamma$ we define the context extension $\ctxext{\Gamma}A$ to be
$\sm{i:\Gamma}A(i)$. Note that context extension $\ctxext{\Gamma}\blank$
can be seen acting not only as a function from the context of the model
$\Gamma\to\UU$ to the context of the model $\UU$, but it also acts on the types 
and terms: when $A$ is a type in context $\Gamma$ we may take the identity map
from $\mftyp{\Gamma\to\UU}{A}\to\mftyp{\UU}{\ctxext{\Gamma}A}$ because
$\mftyp{\Gamma\to\UU}{A}$ is taken to be $\mftyp{\UU}{\ctxext{\Gamma}A}$;
thus, $\ctxext{\Gamma}P\defeq P$ for every family $P$ over $A$ in context 
$\Gamma$. Likewise, when $P$ is a type in context $A$ in the model 
$\Gamma\to\UU$ then context extension should act on the terms of $P$ via a
function $\terms{P}\to\terms{\ctxext{\Gamma}P}$, which we take to be
the identity map once more.

Since $\mftyp{\UU}{\Gamma}$ is a model of type theory it
has it's own notion of context extension: when $P$ is a type in context $A$ in
the model $\mftyp{\UU}{\Gamma}$ then $\ctxext{A}P$ is the family
$\lam{i}\sm{x:A(i)}P(i,x):\Gamma\to\UU$. Also the context extension of
$\Gamma\to\UU$ acts trivially on the types and terms. Context extension is
analoguous to the Grothendieck construction that associates the category of
elements to a presheaf and it gives us $\Sigma$-types.

@@

We will interpret the dependent product $\mprd{A}{P}$ of a family $P$ over
$A$ in context $\Gamma$ by
\begin{equation*}
\mprd{A}{P}(i)\defeq \prd{x:A(i)}P(i,x)
\end{equation*}
With this interpretation there is an equivalence 
$\lambda:\eqv{\terms{P}}{\terms{\mprd{A}{P}}}$. This is lambda-abstraction; its
inverse being evaluation. We should note however, that the rule for evaluation
we interpret here is
\begin{equation*}
\inference{\Gamma\vdash f:\mprd{A}{P}}{\ctxext{\Gamma}A\vdash\tfev(f):P}
\end{equation*}
which is different than the usual rule
\begin{equation*}
\inference{\Gamma\vdash f:\mprd{A}{P} \qquad \Gamma\vdash a:A}{\Gamma \vdash \tfev(f,a):\subst{a}{P}}
\end{equation*}
The reason for this is that the interpretation of the usual rule would give a
function of type $\terms{\mprd{A}{P}}\to\prd{x:\terms{A}}\terms{\subst{x}{P}}$,
but this does not describe the terms of $\mprd{A}{P}$ by any means. Moreover,
since we have implemented substitution, we obtain from $\ctxext{\Gamma}A\vdash\tfev(f):P$ 
and $\Gamma\vdash x:A$ a term $\Gamma\vdash\subst{x}{\tfev(f)}:\subst{x}{P}$ and
therefore we do not loose anything with this approach.

@@

\begin{defn}
Suppose that $\Gamma\vdash A:\mfGraph$. Then we define the graph $\ctxext{\Gamma}A$
by
\begin{align*}
\ctxext{\Gamma}A_0 & \defeq \sm{i:\Gamma_0}A_0(i)\\
\ctxext{\Gamma}A_1(\pairr{i,x},\pairr{j,y}) & \defeq \sm{q:\Gamma_1(i,j)}A_1(q,x,y).
\end{align*}
\end{defn}

\begin{defn}
Suppose that $\Gamma\vdash A:\mfGraph$ and $\Gamma\vdash B:\mfGraph$. Then we
define the graph $\ctxwk{A}{B}$ in context $\ctxext{\Gamma}A$ by
\begin{align*}
(\ctxwk{A}{B})_0(\pairr{i,x}) & \defeq B_0(i)\\
(\ctxwk{A}{B})_1(\pairr{q,e},u,v) & \defeq B_1(q,u,v).
\end{align*}
\end{defn}

@@

\begin{rmk}
We note that the function $\ctxext{\unit}\blank$ is an equivalence
from $\mftyp{\mfGraph}{\unit}$ to $\tfGraph$. It's inverse is the function which maps
a graph $\Gamma$ to the pair $\pairr{A_0,A_1}$ where $A_0$ is defined by
$A_0(\ttt)\defeq\Gamma_0$ and where $A_1$ is defined by $A_1(\ttt)\defeq
\Gamma_1$.

@@

\begin{defn}
Suppose that $\Gamma\vdash A:\mfGraph$. Then we define the graph $\ctxext{\Gamma}A$
by
\begin{align*}
(\ctxext{\Gamma}A)_0 & \defeq \sm{i:\Gamma_0}A_0(i)\\
(\ctxext{\Gamma}A)_1(\pairr{i,x},\pairr{j,y}) & \defeq \sm{q:\Gamma_1(i,j)}A_1(q,x,y).
\end{align*}
There is a morphism $\pi^A:\ctxhom{\ctxext{\Gamma}A,\Gamma}$ given by
\begin{align*}
\pi^A_0 & \defeq \lam{\pairr{i,x}}i && : (\sm{i:\Gamma_0}A_0(i))\to \Gamma_0\\
\pi^A_1(\pairr{i,x},\pairr{j,y}) & \defeq \lam{\pairr{q,e}}q && : (\sm{q:\Gamma_1(i,j)}A_1(q,x,y))\to\Gamma_1(i,j).
\end{align*}
\end{defn}

\begin{rmk}
The construction of $\ctxext{\Gamma}A$ is similar to the Grothendieck
construction of the category of elements for a presheaf.
\end{rmk}

\begin{defn}
Let $\Gamma\vdash A:\mfGraph$. A family $P$ of graphs over $A$ is by
definition a graph $P$ in the context $\ctxext{\Gamma}A$. More explicitly, a family
$P$ of graphs over $A$ in context $\Gamma$ consists of
\begin{align*}
P_0 & : \prd*{i:\Gamma_0}A_0(i)\to\type\\
P_1 & : \prd*{i,j:\Gamma_0}{q:\Gamma_1(i,j)}*{x:A_0(i)}*{y:A_0(j)} A_1(q,x,y)\to P_0(x)\to P_0(y)\to\type
%P_1 & : \prd{\pairr{i,x},\pairr{j,y}:(\ctxext{\Gamma}A)_0}(\ctxext{\Gamma}A)_1(\pairr{i,x},\pairr{j,y})\to P_0(i,x)\to P_0(j,y)\to\type.
\end{align*}
\end{defn}

@@

\begin{rmk}
For any two graphs $A$ and $B$ in a context $\Gamma$, we have the graph
$B[\pi^A]$ in the context $\ctxext{\Gamma}A$, which we may just write
as $B$. Therefore, we can
consider the graphs $\mprd{A}{B}$ and $\msm{A}{B}$. As usually those shall be
denoted  by $A\to B$ and $A\times B$, respectively.

@@

\begin{defn}
Let $A$ be a graph in context $\Gamma$. We define the family $\idtypevar{A}$ over $\ctxwk{A}{A}$ in
context $\ctxext{\Gamma}A$ by
\begin{align*}
(\idtypevar{A})_0(\pairr{i,x},y) & \defeq \id{x}{y}\\
(\idtypevar{A}){}_1(\pairr{q,e},d,\alpha,\alpha') & \defeq \id{\trans{\pairr{\alpha,\alpha'}}{e}}{d}
\end{align*}
where $q:\Gamma_1(i,j)$, $e:A_1(q,x,x')$, $d:A_1(q,y,y')$, $\alpha:\id{x}{y}$
and $\alpha':\id{x'}{y'}$. The transportation along the path 
$\pairr{\alpha,\alpha'}:\id{\pairr{x,x'}}{\pairr{y,y'}}$ in $A_0(i)\times A_0(j)$
is taken with respect to the family $\lam{x}{x'}A_1(q,x,x')$.

@@

\begin{defn}
Let $D$ be a family over $\idtypevar{A}$ in context 
$\ctxext({\Gamma}{A})\ctxwk{A}{A}$. Then we have the family
$\subst{\idfunc[A]}{D}$ over $\subst{\idfunc[A]}{\idtypevar{A}}$ in
context $\ctxext{\Gamma}A$ given by
\begin{align*}
\subst{\idfunc[A]}{D}_0(\pairr{i,x},\alpha) & \defeq D_0(\pairr{i,x,x},\alpha)\\
\subst{\idfunc[A]}{D}_1(\pairr{q,e},\gamma) & \defeq D_1(\pairr{q,e,e},\gamma).
\end{align*}
The family $\subst{\reflf{A}}{\subst{\idfunc[A]}{D}}$ over $A$ in
context $\Gamma$ is given by
\begin{align*}
\subst{\reflf{A}}{\subst{\idfunc[A]}{D}}_0(i,x) & \defeq D_0(i,x,x,\refl{x})\\
\subst{\reflf{A}}{\subst{\idfunc[A]}{D}}_1(q,e) & \defeq D_1(q,e,e,\refl{e}).
\end{align*}
To show that the identity graphs correctly interpret the identity elimination
rule, we must give a function
\begin{equation*}
\tfJ : \terms{\subst{\reflf{A}}{\subst{\idfunc[A]}{D}}}\to\terms{D}.
\end{equation*}
Note that a term $d$ of $\subst{\reflf{A}}{\subst{\idfunc[A]}{D}}$
consists of
\begin{align*}
d_0 & : \prd*{i:\Gamma_0}{x:A_0(i)}D_0(i,x,x,\refl{x})\\
d_1 & : \prd*{i,j:\Gamma_0}{q:\Gamma_1(i,j)}*{x:A_0(i)}*{y:A_0(j)}{e:A_1(q,x,y)}D_1(q,e,e,\refl{e})
\end{align*}
A simple argument using path induction reveals that terms of
$\subst{\reflf{A}}{\subst{\idfunc[A]}{D}}$ indeed yield terms of $D$. 
\end{defn}

@@

\begin{lem}
Let $A$ and $B$ be graphs in a context $\Gamma$ and let $f:A\to B$. The following are equivalent:
\begin{enumerate}
\item $f[i]:A[i]\to B[i]$ is an equivalence of graphs for every term $i:\Gamma$.
\item $\ctxext{\Gamma}f:\ctxext{\Gamma}A\to\ctxext{\Gamma}B$ is an equivalence of graphs.
\item Both $f_0(i)$ and $f_1(q,x,y)$ are always equivalences.
\item $\terms{f}:\terms{\Delta}\to\terms{\Gamma}$ is an equivalence.
\end{enumerate}
\end{lem}

\begin{rmk}
It follows that there is an equivalence
\begin{equation*}
\ctxext{\Gamma}\msm{A}{P}\simeq\ctxext({\Gamma}{A})P
\end{equation*}
for every family $P$ of graphs over a graph $A$ in context $\Gamma$.
\end{rmk}

>>>>

When $A$ is a type in context
$\Gamma$ we define the context extension $\ctxext{\Gamma}{A}$ to be
$\sm{i:\Gamma}A(i)$. Note that context extension $\ctxext{\Gamma}{\blank}$
can be seen acting not only as a function from the context of the model
$\Gamma\to\UU$ to the context of the model $\UU$, but it also acts on the types 
and terms: when $A$ is a type in context $\Gamma$ we may take the identity map
from $\mftyp{\Gamma\to\UU}{A}\to\mftyp{\UU}{\ctxext{\Gamma}{A}}$ because
$\mftyp{\Gamma\to\UU}{A}$ is taken to be $\mftyp{\UU}{\ctxext{\Gamma}{A}}$;
thus, $\ctxext{\Gamma}{P}\defeq P$ for every family $P$ over $A$ in context 
$\Gamma$. Likewise, when $P$ is a type in context $A$ in the model 
$\Gamma\to\UU$ then context extension should act on the terms of $P$ via a
function $\terms{P}\to\terms{\ctxext{\Gamma}{P}}$, which we take to be
the identity map once more.

Since $\mftyp{\UU}{\Gamma}$ is a model of type theory it
has it's own notion of context extension: when $P$ is a type in context $A$ in
the model $\mftyp{\UU}{\Gamma}$ then $\ctxext{A}{P}$ is the family
$\lam{i}\sm{x:A(i)}P(i,x):\Gamma\to\UU$. Also the context extension of
$\Gamma\to\UU$ acts trivially on the types and terms. Context extension is
analoguous to the Grothendieck construction that associates the category of
elements to a presheaf and it gives us $\Sigma$-types.

@@

We will interpret the dependent product $\mprd{A}{P}$ of a family $P$ over
$A$ in context $\Gamma$ by
\begin{equation*}
\mprd{A}{P}(i)\defeq \prd{x:A(i)}P(i,x)
\end{equation*}
With this interpretation there is an equivalence 
$\lambda:\eqv{\terms{P}}{\terms{\mprd{A}{P}}}$. This is lambda-abstraction; its
inverse being evaluation. We should note however, that the rule for evaluation
we interpret here is
\begin{equation*}
\inference{\Gamma\vdash f:\mprd{A}{P}}{\ctxext{\Gamma}{A}\vdash\tfev(f):P}
\end{equation*}
which is different than the usual rule
\begin{equation*}
\inference{\Gamma\vdash f:\mprd{A}{P} \qquad \Gamma\vdash a:A}{\Gamma \vdash \tfev(f,a):\subst{a}{P}}
\end{equation*}
The reason for this is that the interpretation of the usual rule would give a
function of type $\terms{\mprd{A}{P}}\to\prd{x:\terms{A}}\terms{\subst{x}{P}}$,
but this does not describe the terms of $\mprd{A}{P}$ by any means. Moreover,
since we have implemented substitution, we obtain from $\ctxext{\Gamma}{A}\vdash\tfev(f):P$ 
and $\Gamma\vdash x:A$ a term $\Gamma\vdash\subst{x}{\tfev(f)}:\subst{x}{P}$ and
therefore we do not loose anything with this approach.

@@

\begin{defn}
Suppose that $\Gamma\vdash A:\mfGraph$. Then we define the graph $\ctxext{\Gamma}{A}$
by
\begin{align*}
\ctxext{\Gamma}{A}_0 & \defeq \sm{i:\Gamma_0}A_0(i)\\
\ctxext{\Gamma}{A}_1(\pairr{i,x},\pairr{j,y}) & \defeq \sm{q:\Gamma_1(i,j)}A_1(q,x,y).
\end{align*}
\end{defn}

\begin{defn}
Suppose that $\Gamma\vdash A:\mfGraph$ and $\Gamma\vdash B:\mfGraph$. Then we
define the graph $\ctxwk{A}{B}$ in context $\ctxext{\Gamma}{A}$ by
\begin{align*}
(\ctxwk{A}{B})_0(\pairr{i,x}) & \defeq B_0(i)\\
(\ctxwk{A}{B})_1(\pairr{q,e},u,v) & \defeq B_1(q,u,v).
\end{align*}
\end{defn}

@@

\begin{rmk}
We note that the function $\ctxext{\unit}{\blank}$ is an equivalence
from $\mftyp{\mfGraph}{\unit}$ to $\tfGraph$. It's inverse is the function which maps
a graph $\Gamma$ to the pair $\pairr{A_0,A_1}$ where $A_0$ is defined by
$A_0(\ttt)\defeq\Gamma_0$ and where $A_1$ is defined by $A_1(\ttt)\defeq
\Gamma_1$.

@@

\begin{defn}
Suppose that $\Gamma\vdash A:\mfGraph$. Then we define the graph $\ctxext{\Gamma}{A}$
by
\begin{align*}
(\ctxext{\Gamma}{A})_0 & \defeq \sm{i:\Gamma_0}A_0(i)\\
(\ctxext{\Gamma}{A})_1(\pairr{i,x},\pairr{j,y}) & \defeq \sm{q:\Gamma_1(i,j)}A_1(q,x,y).
\end{align*}
There is a morphism $\pi^A:\ctxhom{\ctxext{\Gamma}{A},\Gamma}$ given by
\begin{align*}
\pi^A_0 & \defeq \lam{\pairr{i,x}}i && : (\sm{i:\Gamma_0}A_0(i))\to \Gamma_0\\
\pi^A_1(\pairr{i,x},\pairr{j,y}) & \defeq \lam{\pairr{q,e}}q && : (\sm{q:\Gamma_1(i,j)}A_1(q,x,y))\to\Gamma_1(i,j).
\end{align*}
\end{defn}

\begin{rmk}
The construction of $\ctxext{\Gamma}{A}$ is similar to the Grothendieck
construction of the category of elements for a presheaf.
\end{rmk}

\begin{defn}
Let $\Gamma\vdash A:\mfGraph$. A family $P$ of graphs over $A$ is by
definition a graph $P$ in the context $\ctxext{\Gamma}{A}$. More explicitly, a family
$P$ of graphs over $A$ in context $\Gamma$ consists of
\begin{align*}
P_0 & : \prd*{i:\Gamma_0}A_0(i)\to\type\\
P_1 & : \prd*{i,j:\Gamma_0}{q:\Gamma_1(i,j)}*{x:A_0(i)}*{y:A_0(j)} A_1(q,x,y)\to P_0(x)\to P_0(y)\to\type
%P_1 & : \prd{\pairr{i,x},\pairr{j,y}:(\ctxext{\Gamma}{A})_0}(\ctxext{\Gamma}{A})_1(\pairr{i,x},\pairr{j,y})\to P_0(i,x)\to P_0(j,y)\to\type.
\end{align*}
\end{defn}

@@

\begin{rmk}
For any two graphs $A$ and $B$ in a context $\Gamma$, we have the graph
$B[\pi^A]$ in the context $\ctxext{\Gamma}{A}$, which we may just write
as $B$. Therefore, we can
consider the graphs $\mprd{A}{B}$ and $\msm{A}{B}$. As usually those shall be
denoted  by $A\to B$ and $A\times B$, respectively.

@@

\begin{defn}
Let $A$ be a graph in context $\Gamma$. We define the family $\idtypevar{A}$ over $\ctxwk{A}{A}$ in
context $\ctxext{\Gamma}{A}$ by
\begin{align*}
(\idtypevar{A})_0(\pairr{i,x},y) & \defeq \id{x}{y}\\
(\idtypevar{A}){}_1(\pairr{q,e},d,\alpha,\alpha') & \defeq \id{\trans{\pairr{\alpha,\alpha'}}{e}}{d}
\end{align*}
where $q:\Gamma_1(i,j)$, $e:A_1(q,x,x')$, $d:A_1(q,y,y')$, $\alpha:\id{x}{y}$
and $\alpha':\id{x'}{y'}$. The transportation along the path 
$\pairr{\alpha,\alpha'}:\id{\pairr{x,x'}}{\pairr{y,y'}}$ in $A_0(i)\times A_0(j)$
is taken with respect to the family $\lam{x}{x'}A_1(q,x,x')$.

@@

\begin{defn}
Let $D$ be a family over $\idtypevar{A}$ in context 
$\ctxext{{\Gamma}{A}}{\ctxwk{A}{A}}$. Then we have the family
$\subst{\idfunc[A]}{D}$ over $\subst{\idfunc[A]}{\idtypevar{A}}$ in
context $\ctxext{\Gamma}{A}$ given by
\begin{align*}
\subst{\idfunc[A]}{D}_0(\pairr{i,x},\alpha) & \defeq D_0(\pairr{i,x,x},\alpha)\\
\subst{\idfunc[A]}{D}_1(\pairr{q,e},\gamma) & \defeq D_1(\pairr{q,e,e},\gamma).
\end{align*}
The family $\subst{\reflf{A}}{\subst{\idfunc[A]}{D}}$ over $A$ in
context $\Gamma$ is given by
\begin{align*}
\subst{\reflf{A}}{\subst{\idfunc[A]}{D}}_0(i,x) & \defeq D_0(i,x,x,\refl{x})\\
\subst{\reflf{A}}{\subst{\idfunc[A]}{D}}_1(q,e) & \defeq D_1(q,e,e,\refl{e}).
\end{align*}
To show that the identity graphs correctly interpret the identity elimination
rule, we must give a function
\begin{equation*}
\tfJ : \terms{\subst{\reflf{A}}{\subst{\idfunc[A]}{D}}}\to\terms{D}.
\end{equation*}
Note that a term $d$ of $\subst{\reflf{A}}{\subst{\idfunc[A]}{D}}$
consists of
\begin{align*}
d_0 & : \prd*{i:\Gamma_0}{x:A_0(i)}D_0(i,x,x,\refl{x})\\
d_1 & : \prd*{i,j:\Gamma_0}{q:\Gamma_1(i,j)}*{x:A_0(i)}*{y:A_0(j)}{e:A_1(q,x,y)}D_1(q,e,e,\refl{e})
\end{align*}
A simple argument using path induction reveals that terms of
$\subst{\reflf{A}}{\subst{\idfunc[A]}{D}}$ indeed yield terms of $D$. 
\end{defn}

@@

\begin{lem}
Let $A$ and $B$ be graphs in a context $\Gamma$ and let $f:A\to B$. The following are equivalent:
\begin{enumerate}
\item $f[i]:A[i]\to B[i]$ is an equivalence of graphs for every term $i:\Gamma$.
\item $\ctxext{\Gamma}{f}:\ctxext{\Gamma}{A}\to\ctxext{\Gamma}{B}$ is an equivalence of graphs.
\item Both $f_0(i)$ and $f_1(q,x,y)$ are always equivalences.
\item $\terms{f}:\terms{\Delta}\to\terms{\Gamma}$ is an equivalence.
\end{enumerate}
\end{lem}

\begin{rmk}
It follows that there is an equivalence
\begin{equation*}
\ctxext{\Gamma}\msm{A}{P}\simeq\ctxext{{\Gamma}{A}}{P}
\end{equation*}
for every family $P$ of graphs over a graph $A$ in context $\Gamma$.
\end{rmk}

### 2013-10-24 article-models-intro.tex
putting the new macros \ctxext, \ctxwk and \subst into use

Because we require a \emph{model} of types in a context, all the structure
which we require at the bottom level will be required to exist higher up as well.
Thus, the model of types in a given context $\Gamma$ will have a type of contexts
itself, which can be seen as the type of types in $\Gamma$; it will have its
own notion of types in a context, its own notion of terms, context extension,
weakening and substitution together with all the structure require for it. For
instance, when $A$ is a type in context $\Gamma$ in a model $\mfM$, then there
is the model of types in context $A$, which is the model of families over $A$. 
This model is required to be \emph{definitionally equal to} the model of types
in the context $\ctxext{\Gamma}A$, the context extension of $\Gamma$ and $A$.
In this way we protect ourselves from the need to dig an infinitely deep structure
of models when we want to consider examples.

>>>>

Because we require a \emph{model} of types in a context, all the structure
which we require at the bottom level will be required to exist higher up as well.
Thus, the model of types in a given context $\Gamma$ will have a type of contexts
itself, which can be seen as the type of types in $\Gamma$; it will have its
own notion of types in a context, its own notion of terms, context extension,
weakening and substitution together with all the structure require for it. For
instance, when $A$ is a type in context $\Gamma$ in a model $\mfM$, then there
is the model of types in context $A$, which is the model of families over $A$. 
This model is required to be \emph{definitionally equal to} the model of types
in the context $\ctxext{\Gamma}{A}$, the context extension of $\Gamma$ and $A$.
In this way we protect ourselves from the need to dig an infinitely deep structure
of models when we want to consider examples.

### 2013-10-24 article-models-main.tex
resolve issue #7
putting the new macros \ctxext, \ctxwk and \subst into use

\begin{comment}
\section*{Introduction\phantomsection\addcontentsline{toc}{section}{Introduction}}
This paper contributes to the univalent homotopy type theory~\cite{kapulkin2012univalence,awodey2012type,pelayo2012homotopy}
program which connects and extends a number of topics such as type theory, (elementary) topos
theory~\cite{MacLaneMoerdijk,johnstone:elephant}, homotopy theory, higher category theory and higher topos
theory~\cite{rezk2010toposes,lurie2009higher}.
A key missing ingredient is a first-order (`elementary') definition of a higher topos. It has been
conjectured~\cite{awodey2012type,shulman2012univalence} that homotopy type theory can be used as the internal language, and thus may
provide an elementary definition of a higher topos; see also~\cite{LumsdaineWarren}.

A main ingredient in the theory of higher toposes is the descent property, which
is described in~\cite{rezk2010toposes,lurie2009higher}. The (weak) descent property
of an ordinary topos $\mathcal{E}$ is the property that
\begin{enumerate}
\item For any family $\{X_i\}_{i\in I}$ of objects of $\mathcal{E}$ with coproduct
$X:=\coprod_{i\in I}X_i$ and any function $f:Y\to X$ we get an isomorphism
\begin{equation*}
Y\cong\coprod_{i\in I} Y\times_X X_i
\end{equation*}
given by the canonical map.
\item Conversely, for any family $\{f_i:X_i\to Y_i\}_{i\in I}$ of maps inducing
$f:\coprod_{i\in I}X_i\to\coprod_{i\in I}Y_i$, the canonical map determines an
isomporphism
\begin{equation*}
Y_i\cong Y\times_X X_i.
\end{equation*}
\item
\item
\end{enumerate}
In model toposes, the descent property can be strengthened considerably, but
ordinary limits and colimits are to be replaced by \emph{homotopy} limits and
colimits. Since the only notion of limits and colimits in type theory is that
of homotopy limit and colimit, this is no restriction for us. The descent
property of model toposes as formulated by Rezk~\cite{rezk2010toposes} is: Let
$X:\mathbf{I}\to\mathcal{E}$ be a functor from a small category into a
model topos $\mathcal{E}$, with homotopy colimit $\bar{X}$. Then we have
\begin{enumerate}
\item Let $\bar{Y}$ be an object and let $f:\bar{Y}\to\bar{X}$ 
be any map in $\mathcal{E}$. Define
the functor $Y:\mathbf{I}\to\mathcal{E}$ by $Y(i):=Y\times_{\bar{X}} X_i$. Then
the canonical map $\typefont{hocolim}(Y)\to\bar{Y}$ is a weak equivalence.
\item Let $Y:\mathbf{I}\to\mathcal{E}$ be a functor and let $f:Y\to X$ be an
equifibered transformation from $Y$ to $X$. Define $\bar{f}:\bar{Y}\to\bar{X}$
be the homotopy colimit of $f$. Then the canonical map 
$Y_i\to \bar{Y}\times^h_{\bar{X}} X_i$ is a weak equivalence.
\end{enumerate}
These two properties are two sides of the same coin and in our type theoretical
setting it is natural to stress this. Given a diagram $D$, we will be able
to summarize the entire descent property in an equivalence between the type
$\typefont{equiFib}(D)$ of equifibered diagrams over $D$ and the type of functions
into $\typefont{colim}(D)$. We prove the descent property in 
theorem~\ref{thm:descent}.

We use universes ala Russell, universe polymorphism and typical ambiguity and hence create the illusion that $\type:\type$;
see~\cite{TheBook}. There are only a few places where we need to be explicit about universe levels.
As is common in HoTT, we will freely use the axiom of function extensionality.
Using this axiom, all limits can be constructed in type theory~\cite{Avigad:limits}.
We use higher inductive types (HITs) to construct co-limits; see~\cite{LumsdaineShulman,Shulman:HIT}.
Using the univalence axiom one can show that $\type$ is an object classifier~\cite{RijkeSpitters:Sets}.
An important feature property of $\infty$-toposes is the descent property. We prove this property in theorem~\ref{thm:descent}.

In the previous paragraph we have used informal categorical terminology such as `limits' to refer to what would be interpreted as homotopy
limits in the $\infty$-category of $\infty$-groupoids. We will continue to use this naive $\infty$-category style without any rigorous
claims to their interpretation in the model.

The paper is organized as follows. We start by defining limits and colimits of over graph. This is the most general case, as we
do not include the possibility of including higher coherence conditions.
We introduce diagrams and (co)limits in section~\ref{sec:diagram}. In section~\ref{sec:suspensions} we study suspensions.
Functions with contractible mapping cones are the same as epimorphisms, as we will see in section~\ref{sec:mapping}.
Finally, we prove the descent property.

\subsection{Notation}
A typical way we write a type in this article is:
\begin{equation}\label{eq:notation_example}
\prd{x,y:A}{p:B(x,y)}C(x,y,p)\to D(x,y,p).
\end{equation}
Assuming that $A$, $B$, $C$ and $D$ have been introduced, this type denotes a
dependent product, where the product ranges over $x,y:A$ and $p:B(x,y)$ and
the factors are the (non-dependent) function types $C(x,y,p)\to D(x,y,p)$. A
more common notation for the same type in mathematics would be
\begin{equation*}
\prod_{x,y:A}\prod_{p:B(x,y)}C(x,y,p)\to D(x,y,p),
\end{equation*}
like a big product of function spaces. Nevertheless we have chosen to use this
less common notation, because the types $A$ or $B$ can become quite complex.
% and formulas would become messy as as complex information starts `dancing
% around'.

Another thing we notice in formula \eqref{eq:notation_example} is the use of
curly braces around $x,y:A$. This indicates that whenever we have a term
\begin{equation*}
f:\prd{x,y:A}{p:B(x,y)}C(x,y,p)\to D(x,y,p),
\end{equation*}
we will omit reference to $x$ and $y$ whenever $f$ is applied to $x,y:A$ and
$p:B(x,y)$. In other words, we will write down $f(p):C(x,y,p)\to D(x,y,p)$
instead of $f(x,y,p)$. We will do this whenever it is possible to retrieve
the terms $x$ and $y$ from the term $p$, and in many similar situations like
this. Back to formula \autoref{eq:notation_example}, it is quite possible 
that we would have used the curly braces in the introduction
of $C$ and $D$ as well, and if we had done that we would have written $C(p)$
and $D(p)$.
\end{comment}

>>>>

\input{article-models-intro}

\input{article-models-eg}
